Tabular Constraint Learning

### 2016-04-13 Paper/ECAI/tabular.tex
Put Tias comments in
editing the paper

\usepackage[dvipsnames]{xcolor}
\newcommand{\sergey}[1]{\textcolor{magenta}{{\sc Sergey:} #1}\xspace}
\newcommand{\samuel}[1]{\textcolor{green}{{\sc Samuel:} #1}\xspace}

@@

\begin{abstract}
  abstract
\end{abstract}
\section{Introduction}
Imagine you have received a CSV file. Upon examining the content it becomes fairly clear it was made in Excel. Even the names of the fields such as Total or Avg indicated the usage of the standard functions. Having a closer look, you could deduce the functional dependencies in the table that look as in Figure \ref{fig:main_example}. Provided that a spreadsheet and the numbers in it are small, and dependencies are not complicated, one can examine the spreadsheet and find all connections. But wouldn't it be good to do it automatically?

This brings us to the \textbf{key question}:\\
Can we discover or reconstruct structural relations in flat tabular spreadsheet data? [in a general way that allows declarative specification of constraints to discover]\\

Of course, it is stated semi-informally here, but it already allows us to provide an intuition on the task. We know that often people work in Excel with ranges, e.g. if we look at the Table 1  and at the first row in the Table 2 in Figure \ref{fig:main_example}, it is clear that one could have selected the first columns, clicked on the \textit{SUM} function and then just used the drag fill handle to apply the formula to the other columns. Similarly, that would work for the second (\textit{Average}), third (\textit{Min})  and forth (\textit{Max}) rows in Table 2. I.e., all operations are vectorized and make use of the standard Excel functions.

@@

\sergey{we need to incorporate the motivation part like error checking and autocompletion in the intro}

@@

@@

\section{Formalization}

\samuel{Explain concepts on example}
\subsection{Groups, Type-consistency and Constraints}
In this work we distinguish the following types of data: numeric and textual. The numeric type has two subtypes: integers and floats. We also consider the special element called \textit{None}, which has two types: numeric and textual. A set is called \textit{type-consistent} iff all elements are either numeric or textual. Certain constraints, such as \textit{rank} or \textit{series}, make use of the numeric sub-types by requiring its arguments to be integers.

@@

An \textit{Excel constraint} is a triple \textit{(\CName, \CSignature, \CFunction)}. Let us elaborate on each of them. \textit{\CName} is the textual name of the constraint together with its variable names. \textit{\CSignature} is a set of constraints specifying the properties of the group assignments, corresponding to the variables in \CName, such as their types, e.g., requiring them to be integers or constraining the sizes, e.g. the length of vectors in the arguments must be equal. These are constraints on the group meta-information not on the actual group content. \textit{\CFunction} is a set of constraints specifying that the data in the subgroups satisfies the function. For example, an excel constraint \textit{rank} has \CName \textit{Y = RANK(X)}, where $X$ and $Y$ are the variables; its \CSignature is: the group $G_X$ (associated with X) is numeric, $G_Y$ is integer and the length of vectors in $G_X$ is the same as in $G_Y$; its \CFunction is the following constraint: a pair of vectors $X,Y$ is a solution iff $X \in G_X, Y \in G_Y$ and each value in $X$ has the rank (possibly with ties) specified in $Y$.

@@

  \end{tabularx}
  \caption{Excel Tabular Constraints \sergey{Luc, we need your comments here on the notation of $X,Y$ and $G_X,G_Y$}}
  \label{table:constraints}
\end{table*}

\samuel{Move to beginning without algo}
% \subsection{Workflow}
% \begin{algorithm}[thb]
%   \begin{algorithmic}
%     \footnotesize
%     \State \textbf{Input:} $D$ -- dataset, \constraints -- constraints, \dependencies -- dependencies \\(optional: tables $T$, groups $G$)
%     \State \textbf{Output:} $S$ -- learned constraints with their satisfaction assignment
%     \If{$T$ is \textbf{not} provided}
%       \State $T \gets \extracttables(D)$
%     \EndIf
%     \If{$G$ is \textbf{not} provided}
%       \State $G \gets \extractgroups(D, T)$
%     \EndIf
%     \State $S \gets \learnconstraints(G,\constraints,\dependencies)$
%     \State \Return $S$
% \end{algorithmic}
% \caption{Workflow}
% \label{algo:workflow}
% \end{algorithm}

>>>>

\usepackage[dvipsnames]{xcolor}
\newcommand{\sergey}[1]{\textcolor{magenta}{{\sc Sergey:} #1}\xspace}
\newcommand{\samuel}[1]{\textcolor{green}{{\sc Samuel:} #1}\xspace}
\newcommand{\tias}[1]{\textcolor{blue}{{\sc Tias:} #1}\xspace}

@@

\begin{abstract}
  abstract
\end{abstract}

\tias{Use spreadsheet instead of Excel}
\tias{Spreadsheet is an instance of tabular data, but not the only one?}

\tias{Rewrite}
\section{Introduction}
Millions of people across use spreadsheets everyday. Yet, many people lack a complete understanding of the structures and existing dependencies in the data. Indeed, the most useful part of spreadsheets is often the formulas and often these formula disappear, when a file is exported and shared as a CSV file. This brings us to the question: can we discover or reconstruct structural relations in flat tabular spreadsheet data? [in a general way that allows declarative specification of constraints to discover]

Let us look at an example in Figure \ref{fig:main_example}, the name already suggest the usage of standard spreadsheet operations such as \textit{average} or \textit{sum}. If we look at the Table 1 and at the first row in the Table 2, it is clear that one could have selected the column called \textbf{1st Quarter}, clicked on the \textit{SUM} function and then just used the drag fill handle to apply the formula to the other columns. Similarly, that would work for the second (\textit{Average}), third (\textit{Min})  and forth (\textit{Max}) rows in Table 2. I.e., all operations are vectorized and make use of the standard spreadsheet operations.

@@

This makes tabular data understanding and constraint learning new and challenging, and it is not only challenging for a machine but for a human as well. Due to the complexity of numeric computations in spreadsheets, people often fail to grasp the properties of the data, which leads to so called the Spreadsheet risk\footnote{\url{https://en.wikipedia.org/wiki/Spreadsheet\#Spreadsheet_risk}} and even caused major flaw in the famous economics papers \cite{flaw_excel} and caused billion-losses in the financial industry. Tabular constraint learning provides a potential cure by indicating learned constraint violations and suggest possible fixes.

@@

\samuel{Spreadsheet risks: financial losses due to errors in spreadsheets}

@@

\section{Formalization}
\tias{More gentle}
\samuel{Explain concepts on example}
\subsection{Groups, Type-consistency and Constraints}
In this work we distinguish the following types of data: numeric and textual. The numeric type has two subtypes: integers and floats. We also consider the special element called \textit{None}, which has two types: numeric and textual. A set is called \textit{type-consistent} iff all elements are either numeric or textual. Certain constraints, such as \textit{rank} or \textit{series}, make use of the numeric sub-types by requiring its arguments to be integers.

@@

\tias{Use generic constraint}
An \textit{Excel constraint} is a triple \textit{(\CName, \CSignature, \CFunction)}. Let us elaborate on each of them. \textit{\CName} is the textual name of the constraint together with its variable names. \textit{\CSignature} is a set of constraints specifying the properties of the group assignments, corresponding to the variables in \CName, such as their types, e.g., requiring them to be integers or constraining the sizes, e.g. the length of vectors in the arguments must be equal. These are constraints on the group meta-information not on the actual group content. \textit{\CFunction} is a set of constraints specifying that the data in the subgroups satisfies the function. For example, an excel constraint \textit{rank} has \CName \textit{Y = RANK(X)}, where $X$ and $Y$ are the variables; its \CSignature is: the group $G_X$ (associated with X) is numeric, $G_Y$ is integer and the length of vectors in $G_X$ is the same as in $G_Y$; its \CFunction is the following constraint: a pair of vectors $X,Y$ is a solution iff $X \in G_X, Y \in G_Y$ and each value in $X$ has the rank (possibly with ties) specified in $Y$.

@@

  \end{tabularx}
  \caption{Excel Tabular Constraints}
  \label{table:constraints}
\end{table*}

% \subsection{Workflow}
% \begin{algorithm}[thb]
%   \begin{algorithmic}
%     \footnotesize
%     \State \textbf{Input:} $D$ -- dataset, \constraints -- constraints, \dependencies -- dependencies \\(optional: tables $T$, groups $G$)
%     \State \textbf{Output:} $S$ -- learned constraints with their satisfaction assignment
%     \If{$T$ is \textbf{not} provided}
%       \State $T \gets \extracttables(D)$
%     \EndIf
%     \If{$G$ is \textbf{not} provided}
%       \State $G \gets \extractgroups(D, T)$
%     \EndIf
%     \State $S \gets \learnconstraints(G,\constraints,\dependencies)$
%     \State \Return $S$
% \end{algorithmic}
% \caption{Workflow}
% \label{algo:workflow}
% \end{algorithm}
